Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-21 paper/verification-conditions.tex
moving things around; result is not entirely coherent

A system is a tuple, $\system = (\inputs, \outputs, \assumes, \guarantees)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, and $\guarantees$ is the set of guarantees.
A system is implemented by a set of components, $\components$, with their various connections.
A component $\component$ is defined by its own tuple in a manner similar to the system as $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component)$.
From the components and connections, $\mathbb{I}_\component$ is defined such that it is the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined such that it is the set of components that provide the output for the system.

An system is correct if and only if for all $\pi$ and for all $i \ge 0$ the following holds:
\[
\begin{array}{lll}
        & \forall \component \in \components &  
            \globally(\historically(\assumes \wedge 
            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime}) 
            \implies \assumes_\component) \\
 \wedge &   & 
            \globally(\historically(\assumes \wedge 
            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime}) 
            \implies \guarantees)
\end{array}
\]
The first condition verifies the input assumptions on each component under the system assumptions and upstream component guarantees. 
It checks if the component guarantees and system assumptions are strong enough to imply input assumptions on all immediate downstream components.
The second condition checks the output guarantees of the system under the system assumptions and component guarantees that provide the output. 
It checks if the guarantees on components providing primary outputs are strong enough to imply the system guarantees.

@@

The expanded property lists in \figref{fig:example-certificate} and \figref{fig:hardened-certificate} are the results from disproving or proving out the above verification conditions.
The additional unexpanded results at the bottom of the figures prove \emph{self-consistency} in the contracts.
It is not uncommon to accidentally write contracts that are self-contradicting.
For example, a contract may guarantee an output be two different values in the same moment of time.
AGREE generates additional verification conditions that prove each component contract, and the composition of contracts, self-consistent.

>>>>

A \emph{system} $\system = (\inputs, \outputs, \assumes,
\guarantees)$, where $\inputs$ is the input set, $\outputs$ is the
output set, $\assumes$ is the set of assumptions, and $\guarantees$ is
the set of guarantees.  A system is implemented by a set of
\emph{components}, $\components$, together with with their
inter-connections.  A component $\component$ is defined by its own
tuple in a manner similar to the system as $(\inputs_\component,
\outputs_\component, \assumes_\component, \guarantees_\component)$.
From the components and connections, $\mathbb{I}_\component$ is
defined to be the set of components providing input to some component
$\component$ in the system, and $\mathbb{O}$ is defined to be the set
of components that provide the output for the system.  A system is
correct if and only if for all $\pi$ and for all $i \ge 0$ the
following conditions hold:
\[
\begin{array}{lll}
        & \forall \component \in \components &
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime})
            \implies \assumes_\component) \\
 \wedge &   &
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime})
            \implies \guarantees)
\end{array}
\]
The first condition verifies the input assumptions on each component
under the system assumptions and upstream component guarantees.  It
checks if the component guarantees and system assumptions are strong
enough to imply input assumptions on all immediate downstream
components.  The second condition checks the output guarantees of the
system under the system assumptions and component guarantees that
provide the output.  It checks if the guarantees on components
providing primary outputs are strong enough to imply the system
guarantees.

@@

The expanded property lists in \figref{fig:example-certificate} and
\figref{fig:hardened-certificate} are the results from verifying or
disproving the above verification conditions.  The additional
unexpanded results at the bottom of the figures prove
\emph{self-consistency} in the contracts.  It is not uncommon to
accidentally write contracts that are self-contradicting.  For
example, a contract may guarantee an output be two different values in
the same moment of time.  AGREE generates additional verification
conditions that prove each component contract, and the composition of
contracts, self-consistent.\footnote{KLS:Is this the same as
satisfiable?}

\subsection{Syntax and semantics of AGREE}
\label{agree-semantics}

The syntax of AGREE is essentially that of quantifier-free first order
logic (FOL) supplemented with a few temporal operators. The terms
(\emph{e}) are the usual arithmetic expressions built from variables
($v$) and numeric and boolean literals ($c$), while formulas
(\emph{b}) are built using the standard logical connectives from
atomic formulas (\emph{a}) based on the usual comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

Supplementing this basic syntax are is a conditional,
$\itelse{b}{(-)}{(-)}$, for both terms and formulas. There are also
temporal operators $\konst{pre}(-)$, delay $(-) \to (-)$, and
$\konst{Hist}(-)$.

The semantics of terms and formulas is in terms of \emph{streams of
values}. Values encompass at least booleans and numbers, but can be
readily extended to include records and arrays. A value stream is a
total function from time (natural numbers) to values:
\[
 \konst{stream} = \mathbb{N} \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}^E_t$
of terms and formulas, omitting the temporal operators, is
conventional. The value of a variable $v$ at time $t$ is found by
looking up the stream bound to $v$ in $E$ (call it $s$) and extracting
the value of $s$ at time $t$.

\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

The temporal operators deal with time in more significant ways. The
value of $\konst{pre}(e)$ at time $t$ is the value of $e$ at time
$t-1$ (at time zero, \konst{pre} is undefined).  A delay $e_1 \to e_2$
temporally ``shifts'' $e_2$ by means of prepending the first element
of $e_1$ to it.

\[
\begin{array}{rcl}
\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
\end{array}
\]

Although basic, these definitions can be used to define higher-level
operators from PTLTL, such as \konst{Once} and \konst{Since}.

\subsection{Code contracts}

Generally, AGREE specifications do not describe the computation
that a component performs. This is entirely by design: AGREE is
intended to reason about component behavior solely at the
specification level. However, the syntax of AGREE specifications
provides enough expressiveness to support the notion of a
\emph{code contract}: a component contract from which an implementation
can be extracted.

\begin{definition}[Output guarantee]
An \emph{output guarantee} is a stylized guarantee that fully
specifies the data written to an output port. There are three
possibilities according to whether the output port $p$ is
a \konst{data} port, an \konst{event} port, or an \konst{event data}
port:
\[
\begin{array}{ll}
\konst{data}: &  p = \mathit{e} \\
\konst{event}: &  \konst{event} (p) = \mathit{b} \\
\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
\end{array}
\]
\end{definition}

\begin{definition}[Code contract]
A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
if $\mathit{Eqs} \cup G \subseteq P_c$, where
\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
of \konst{eq} statements and $G$ is the set of output guarantees, one for
each element of $O_c$. In the interpretation as code, the order of
elements of $\mathit{Eqs}$ is important, and is simply taken to be the
occurrence order of the \konst{eq} statements in the syntax. Thus we
will work with the
\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
\end{definition}
